---
title: "Econometrics II - Problem 9"
author: "William Radaic Peron"
date: \today
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

library(rugarch)
library(ARDL)
library(mFilter)
library(dynlm)
library(tseries)




```


```{r database, include = FALSE}

# Loading the dataframe:

library(readxl)
df <- read_excel("Base de Dados - Problema 9.xlsx")



```

\textbf{Ref:} IPEA Discussion Paper 230, Lecture 8 - Notes, Enders 5.1-5.3.

In this problem, we'll apply ARDL modelling to the Phillips Curve. Namely, we'll focus on two hypothesis for the curve: \begin{itemize}
\item \textbf{Inertial inflation.} There's a backward-looking component to any Phillips curve.
\item \textbf{Rational expectations.} The curve is forward-looking.
\end{itemize}
From these hypothesis, it is also possible to test a hybrid approach that incorporates both a backward and a forward component. 

\section{ARDL models}

Consider the following model: 
$$ Y_t = \alpha + \sum_{j=0}^q \theta_j X_{t-j} + \varepsilon_t $$
This is the *distributed lag* model for $q$ lags of $X$. Note that $Y_t$ has serial correlation because of its relation with $X_{t-j}$.

An *Autoregressive distributed lag* model is a combination of an $AR(p)$ with a $DL(q)$:
$$ Y_t = \alpha + \sum_{i=1}^p \phi_i Y_{t-i} + \sum_{j=0}^q \theta_j X_{t-j} + \varepsilon_t $$
Or, with lag operator notation:
$$ \Phi (L) Y_t = \alpha + \Theta (L) X_t + \varepsilon_t $$ 

We can further define $\Psi := \Phi^{-1} (L) \Theta (L) = \psi_0 + \psi_1 L + ....$ This implies that we can rewrite the ARDL(p,q) model as:
$$ Y_t = \Phi^{-1}(1)\alpha + \Psi (L) X_t + \Phi^{-1} (L) \varepsilon_t $$
Also note that we can expand this definition by letting $X_{t,s}$ be a matrix of $s$ processes.

\section{Phillips Curve and its developments}

Phillips famously postulated a negative relation -- i.e., a trade-off -- between inflation rate and unemployment. We can define this curve as follows:
$$ \pi_t = \alpha - \gamma u_t, $$
where $\pi_t$ is inflation, $u_t$ is unemployment and $\gamma > 0$. Note that this is a *static model*. Furthermore, note that this formulation proposes a *long term trade-off* between these variables.

Friedman and Phelps disputed the claims of these long run effects by including *expectations* in the model. The monetarist Phillips curve asserted that inflation expectations would be formed as a function of *past inflation rates* -- i.e., under the assumption of adaptative expectations. This model has the following representation:
$$ \pi_t = \mathbb{E}_{t-1}(\pi_t) - \gamma (u_t - \bar{u}), $$
where $\bar{u}$ represents the natural rate of unemployment (NAIRU). This is called the Accelerationist Phillips Curve (APC), given its *backward-looking* -- hence, inertial -- nature. Now, the model is *dynamic*, and can be translated to the ARDL framework with $Y_t := \pi_t, X_t := \mathbb{E}_{t-1}(\pi_t)$. 

The rational expectations revolution of Lucas and Sargent, which implied fundamental critiques to the Keynesian system, prompted a revaluation of its framework, including the Phillips Curve. The New Keynesian Phillips Curve (NKPC) incorporates *rational expectations* in the model:
$$\pi_t = \beta \mathbb{E}_t (\pi_{t+1}) + \gamma x_t, $$
where $x_t$ is a measure of output gap. Note that this model is essentially *forward-looking*.  
\begin{quote}
This model implies that there is no need for gradualist policies to reduce inflation. According to the NKPC, low inflation can be achieved immediately by the central bank announcing (and the public believing) that it is committing itself to eliminating positive output gaps in the future. (IPEA, p. 11)
\end{quote}
We can interpret this model in the ARDL framework as a DL model with $X_t := \mathbb{E}_{t}(\pi_{t+1})$.

Some authors have also proposed a NKPC with a backward-looking element. This is called the Hybrid Philips Curve:
$$ \pi_t = \gamma_b \pi_{t-1} + \gamma_f \mathbb{E}_t (\pi_{t+1}) + \kappa x_t. $$
This can be interpreted as an ARDL(1,1) model, with $X_t := \mathbb{E}_{t-1}(\pi_t)$.

\section{Estimation}

These hypothesis will now be tested with Brazilian time series data for inflation, unemployment and inflation expectation.

``` {r ardl estimation}

# APC

nairu <- hpfilter(df$unemployment, 1600, type = 'lambda')

plot(nairu)

nairu_trend <- nairu$trend

u_dev <- df$unemployment - nairu_trend

df$u_dev = u_dev

exp_lag = dplyr::lag(df$exp_IPCA, k=1)

df$exp_lag = exp_lag

df2 = df[2:length(df$IPCA),]

auto_apc <- auto_ardl(IPCA ~ exp_lag + u_dev, data = df2, max_order = 12)

summary(auto_apc$best_model)

Box.test(auto_apc$best_model$residuals)
AIC(auto_apc$best_model)
BIC(auto_apc$best_model)

# APC as DL of order 1

apc1 <- dynlm(IPCA ~ exp_lag + u_dev, data = df2)

summary(apc1)

Box.test(apc1$residuals) # Modelo claramente inconsistente


# NKPC

auto_nkpc <- auto_ardl(IPCA ~ exp_IPCA + u_dev, data = df, max_order = 18)

summary(auto_nkpc$best_model) # Unit root?

adf.test(df2$IPCA) # UNIT ROOT!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

######## Correcting for unit root ############

df3 = data.frame(diff(df$IPCA), diff(df$exp_IPCA), df2$u_dev)

colnames(df3) = c("IPCA", "exp_IPCA", "u_dev")

diff_lag = dplyr::lag(df3$exp_IPCA, 1)

df3$diff_lag = diff_lag

df4 = df3[2:length(df3$IPCA),]

# APC

auto_apc_diff <-  auto_ardl(IPCA ~ diff_lag + u_dev, data = df4, max_order = 24)

summary(auto_apc_diff$best_model)



n = 12
x = matrix(NA, nrow = n, ncol = 1)

for (i in (1:n)) {
  
  auto_apc_diff <- auto_ardl(IPCA ~ diff_lag + u_dev, data = df4, max_order = i)
  
  x[i,] = Box.test(auto_apc_diff$best_model$residuals)$p.value
  
}


# NKPC


auto_nkpc_diff <- auto_ardl(IPCA ~ u_dev | exp_IPCA, data = df3, max_order = 18)

summary(auto_nkpc_diff$best_model)

Box.test(auto_nkpc_diff$best_model$residuals)

nkpc_dyn <- dynlm(IPCA ~ exp_IPCA + u_dev, data = df3)

summary(nkpc_dyn)

Box.test(nkpc_dyn$residuals)

# HPC 

hpc_dyn <- dynlm(IPCA ~ exp_IPCA + L(IPCA, 1) + u_dev, data = df3)

summary(hpc_dyn)

Box.test(hpc_dyn$residuals)

# The results suggest that inflation is both forward and backward looking.


ardl22 <- ardl(IPCA ~ u_dev | exp_IPCA, data = df2, order = c(2,2))

co <- ardl22$coefficients

x <- matrix(NA, nrow = 24, ncol = 1)

x[1,] = co[4]

x[2,] = co[2]*x[1,] + co[5]

x[3,] = co[2]*x[2,] + co[3]*x[1,] + co[6]

for (i in 4:24) {
  x[i,] = co[2]*x[(i-1),] + co[3]*x[(i-2),]
}
 
plot(1:24, x, "l")
sum(x)

ARDL::multipliers(ardl22)

multipliers(ardl22)
```






























