---
title: "Econometrics II - Problem 4"
author: "William Radaic Peron, Vinícius Marcial"
date: \today
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(forecast)
library(AER)
library(tidyverse)
library(tidyr)
library(dplyr)
library(corrplot)
library(ggthemes)
library(gridExtra)
library(gtsummary)
library(readxl)
library(ggthemes)
library(mdscore)
library(readxl)
library(fpp)
library(sidrar)
library(mFilter)
library(detrendr)

```


```{r database, include = FALSE}

# Loading the dataframe:


pib <- get_sidra(6612, variable = 9318, category = 90707, period = "all")

pib_limpo <- pib[(pib$`Setores e subsetores (Código)` == 90707),]

pib <- pib_limpo

pib_limpo2 <- pib[,c(5,13)]

pib <- pib_limpo2

names(pib)[1] <- "t"
names(pib)[2] <- "v"

pib$t <- as.numeric(pib$t)

names(pib)

head(pib)

tail(pib)

# df <- ts(df)


```

In this problem, we'll be forecasting GDP in the short term and creating some models of GDP growth in the long run. This presents some challenges, namely those related to *ergodicity* and *stationarity*. 

``` {r plots}

pplot <- ggplot(data = pib, aes(x = t, y = v)) + geom_line() + ggtitle("Time series plot") + theme_few()
pplot 


```




As we have downloaded the *pure* quarterly data, it presents *seasonality* and an upwards tendency. This implies that the *time series will not be stationary*. Therefore, we need to employ methods that circumvent this issue and assure us that we can continue modelling the series as an ARMA(p,q).

\section{Decomposing the time series}

We will now assume that we can decompose the time series in three distinct elements in an additive model: 
$$ X_t = f_t+ s_t + Y_t $$,
where $f_t$ denotes the tendency of the ts, $s_t$ denotes seasonality, $Y_t$ is stochastic. We also assume that $f_t, s_t$ are *deterministic*.

\subsection{Trend}

First, we'll construct a *parametric* model of the trend. Let's assume that $f_t$ can be modelled by a linear form:
$$ f_t = \gamma_0 + \gamma *t $$

``` {r parametric linear}

linear_trend <- lm(v ~ t, data = pib )

summary(linear_trend)

ggplot(data = pib, aes(x = t, y = v)) + stat_smooth(method = "lm", se = F) + geom_line() + geom_point() + theme_few() + ggtitle("Linear trend, GDP")

```

Another way to find $f_t$ is via a *non-parametric* process. For this, we'll use an HP filter and a moving average.

``` {r hp filter}

pib_ts <- ts(pib$v)

hp_trend <- hpfilter(pib_ts, freq = 1600, type = "lambda")

plot(hp_trend)


```

Now, a moving average.

``` {r moving average}

pib_ma <- ma(pib$v, order = 4)

autoplot(pib_ma, color = "blue") + geom_line(data = pib, aes(x = 1:length(pib$t), y = v), color = "red") + theme_few()

```

\subsection{Seasonality}

We can now create a function for $s_t$. This will be done with dummies:
$$ D_i = 1, i = t $$
$$ D_i = 0 \, otherwise$$ 

``` {r dummies season}

tri <- c(NA)

tri1 <- c(1,2,3,4)

i = 1

while (i < 25) { 

  tri <- append(tri, tri1)
  
  i = i + 1
  
  }

tri <- tri[-1]
tri <- c(tri, 1)

tri

pib <- data.frame(pib, tri)

names(pib)[1] <- "t"
names(pib)[2] <- "v"
names(pib)[3] <- "tri"

dummies <- data.frame(matrix(NA, nrow = length(pib$t), ncol = 4))

for (j in 1:4) {
  
  dummies[j] <- as.numeric(pib$tri == j)
  
}

hp_fitted <- hp_trend[2] 

hp_fitted <- hp_fitted$trend

detrend <- pib$v - hp_fitted

pib <- data.frame(pib, dummies, detrend)

names(pib) <- c("t", "v", "tri", "X1", "X2", "X3", "X4", "detrend")

head(pib)

dummy_lm <- lm(detrend ~ X2 + X3 + X4, data = pib)
  
summary(dummy_lm)

```

\subsection{$Y_t$}

We'll now use the HP-fitered version of $f_t$ and the dummy approach to $s_t$.

``` {r stochastic yt}


yt <- as.vector(pib$v) - (hp_fitted + dummy_lm$fitted.values)

mean(yt)

autoplot(yt) + theme_few()

y <- data.frame(1:97, yt)

names(y) <- c("t", "yt")

y

```



\section{Identifying and estimating ARMA(p,q) for $Y_t$}


We are now in a position to identify and estimate the best model for our time series $Y_t$.

Applying the function *auto.arima* from the package *forecast* to identify and estimate the model:

``` {r estimation autoarima}


aa_model <- auto.arima(y$yt, num.cores = 24, max.d = 0, stepwise = F)

summary(aa_model)

print("t-values: ")

aa_t <- matrix(NA, nrow = aa_model$arma[1] + aa_model$arma[2])

for (i in c(1:(aa_model$arma[1] + aa_model$arma[2]))) { 

aa_t[i] <- aa_model$coef[i]/sqrt(aa_model$var.coef[i,i])

}

aa_t <- data.frame(aa_t)

aa_t

aa_q <- Box.test(aa_model$residuals, lag = aa_model$arma[1] + aa_model$arma[2])
aa_q

criteria <- matrix(NA, nrow = 1, ncol = 3)

aa_criteria <- data.frame("AR(2)*", aa_model$aic, aa_model$bic)

names(aa_criteria) <- c("Model", "AIC", "BIC")

aa_criteria

fac_e <- ggAcf(aa_model$residuals, type = "correlation", lag.max = 30, plot = T) + theme_few()

facp_e <- ggPacf(aa_model$residuals, type = "correlation", lag.max = 30, plot = T) + theme_few()

fac_e
facp_e

mean(aa_model$residuals)

plot(aa_model$residuals)

```

``` {r plots fac facp}


facst <- ggAcf(y$yt, type = "correlation", lag.max = 30, plot = T) + theme_few()
faclt <- ggAcf(y$yt, type = "correlation", lag.max = 5000, plot = T) + theme_few()

facpst <- ggPacf(y$yt, type = "correlation", lag.max = 30, plot = T) + theme_few()
facplt <- ggPacf(y$yt, type = "correlation", lag.max = 5000, plot = T) + theme_few()

facst
faclt

facpst
facplt

```

The results of *auto.arima* imply that the best model is an ARMA(2,0) -- i.e., an AR(2): 
$$ y_t = c + \phi_1 y_{t-1} + \phi_2 y_{t-2} + \varepsilon_t, \hspace{1em} \varepsilon_t \sim wn(0, \sigma^2)$$



``` {r forecast plot }

fc <- forecast(y$yt, model = aa_model, h = 4)

autoplot(fc) + theme_few()


```



\section{Long term GDP growth}


``` {r loading data pme}

unemp <- read_excel("C:/Users/William/Downloads/tabela2176.xlsx")

unemp1 <- as.numeric(unemp[11,])

unemp <- unemp1[-1]

df_unemp <- data.frame(1:length(unemp), unemp)

names(df_unemp) <- c("t", "r")

df_unemp

hp_unemp <- hpfilter(df_unemp$r, freq = 1600, type = "lambda")

plot(hp_unemp)

nairu <- hp_unemp$trend

nairu


```

