---
title: "Econometrics II - Problem 6"
author: "William Radaic Peron, Vinícius Marcial"
date: "3 de setembro de 2020"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

Sys.setenv(RSTUDIO_PANDOC="C:/Users/William/anaconda3/envs/py_r/Library/Scripts/pandoc")

library(ggplot2)
library(forecast)
library(AER)
library(tidyverse)
library(tidyr)
library(dplyr)
library(corrplot)
library(ggthemes)
library(gridExtra)
library(gtsummary)
library(readxl)
library(ggthemes)
library(mdscore)
library(readxl)
library(fpp)
library(sidrar)
library(mFilter)
library(detrendr)
install.packages(c('dynlm','ggfortify','tseries'), repos='http://cran.rstudio.com/')
```

\section{Part 1: Spurious regressions}

{{matéria}}



\section{Part 2: Hypothesis testing for unit roots}

As has been discussed during the lecture, when the population model is a *random walk*:
$$ Y_t = 1*Y_{t-1} + \varepsilon_t, \hspace{1em} \varepsilon \sim wn(0, \sigma^2), $$
it happens that the series *is not ergodic* (nor is it stationary). Therefore, the usual asymptotic properties *do not hold*, as all innovations have *permanent effects*. 
$$ Y_t = c + \delta t + \sum_{i=1}^t \varepsilon_i $$

The random walk, defined above, is an example of the more general class of *unit root processes*:
$$ Y_t = c + \delta t + u_t,$$
$u_t$ has an ARMA(p,q) representation:
$$ \Phi_p (L) u_t = \Theta_q (L) \varepsilon_t, \hspace{1em} \sim wn(0, \sigma^2)$$


``` {r }














```


\section{Part 3: Applying the Dickey-Fuller test for GDP}













