---
title: "Econometrics II - Problem 1"
author: "William Radaic Peron"
date: \today
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
library(ggplot2)
library(forecast)
library(AER)
library(tidyverse)
library(tidyr)
library(dplyr)
library(corrplot)
library(ggthemes)
library(gridExtra)
library(gtsummary)
library(readxl)
library(ggthemes)
library(mdscore)

library(formatR)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

```

Loading the database and creating dummy variables:

```{r database}

df <- read_excel("RS_USD.xlsx")


names(df)[names(df) == "R$/US$"] <- "p"

names(df)[names(df) == "Variação (em %)"] <- "delta"

names(df)[names(df) == "Data"] <- "date"

sign <- as.numeric(df$delta > 0)

count <- c(1:2153)

df <- data.frame(count, df, sign)

```


Before constructing our models, we need to check (intuitively) if the series at hand is *stationary* and *ergodic*. For this, we're going to plot the time series, its autocorrelations and partial autocorrelations.

``` {r plots}

pplot <- ggplot(data = df, aes(x = date, y = p)) + geom_line() + ggtitle("USD/BRL, Price") + theme_few()
pplot 

deltaplot <- ggplot(data = df, aes(x = date, y = delta)) + geom_line() + ggtitle("USD/BRL, %") + theme_few()
deltaplot 

dummyplot <- ggplot(data = df, aes(x = count, y = sign)) + geom_line() + ggtitle("USD/BRL, +/-") + xlim(1,200) + theme_few()
dummyplot

# For delta

acf_delta <- Acf(df$delta, lag.max = 5000)
acf_test_values <- acf_delta$acf/sd(acf_delta$acf)

head(data.frame(acf_test_values))

facst <- ggAcf(df$delta, type = "correlation", lag.max = 20, plot = T) + theme_few()
faclt <- ggAcf(df$delta, type = "correlation", lag.max = 5000, plot = T) + theme_few()

facpst <- ggPacf(df$delta, type = "correlation", lag.max = 100, plot = T) + theme_few()
facplt <- ggPacf(df$delta, type = "correlation", lag.max = 5000, plot = T) + theme_few()

facst
faclt

facpst
facplt

facst2 <- ggAcf((df$delta)^2, type = "correlation", lag.max = 20, plot = T) + theme_few()
facst2

```

Let's now create our first ARMA models (equivalent to ARIMA with 2nd argument = 0). We'll begin with the first hypothesis: $\mathbb{P}(+) = \mathbb{P}(-).$ Modelling this with an AR(1), we have:

$$Sign_{t+1} = \alpha + \beta Sign_t + \varepsilon, \hspace{2em} \varepsilon \sim wn(0, \sigma^2)$$
In R, we'll use the package *forecast* to construct this model:

```{r AR(1)}

AR1sign <- Arima(df$sign, order = c(1, 0, 0))
summary(AR1sign)

plot(AR1sign)

tsdisplay(AR1sign$residuals)


```

With the results of the summary, we can now apply a hypothesis test for our first question.\footnote{Testing $\beta$ is equivalent to testing $\gamma$.}

$$ H_0: \beta = 0$$
$$H_1: \beta \neq 0$$

$\dfrac{\hat{ar_1} - ar_1}{s.e.(ar_1)}$:

``` {r ht 1}

AR1sign$coef[1]/sqrt(AR1sign$var.coef[1,1])


```


The second hypothesis in the problem refers to the delta of the variation: 
$$\mathbb{E}(\Delta | + ) \neq \mathbb{E}(\Delta | - ).$$

$$\Delta_{t+1} = \alpha + \beta Sign_t + \varepsilon, \hspace{2em} \varepsilon \sim wn(0, \sigma^2).$$
``` {r lm signt}

lmsignt <- lm(delta ~ lag(df$sign, k = 1), data = df)
summary(lmsignt)

ggplot(df, aes(x = lag(df$sign, k = 1), y = delta)) + geom_smooth(method = "lm")

```


$$\Delta_{t+1} = \alpha + \beta_1 \Delta_t + \beta_2Sign_t + \varepsilon, \hspace{2em} \varepsilon \sim wn(0, \sigma^2)$$

``` {r delta 1}

AR1delta <- Arima(df$delta, order = c(1, 0, 0), xreg = lag(df$sign, k = 1))
summary(AR1delta)

AR1delta$coef[1]/sqrt(AR1delta$var.coef[1,1])

plot(AR1delta)

tsdisplay(AR1delta$residuals)




```



The last hypothesis in the problem refers to the variance: 
$$\mathbb{E}(\Delta_{t+1}^2 | \Delta_t).$$

$$\Delta^2_{t+1} = \alpha + \beta \Delta^2_t + \varepsilon, \hspace{2em} \varepsilon \sim wn(0, \sigma^2)$$

``` {r var1}

AR1var <- Arima((df$delta)^2, order = c(1, 0, 0))
summary(AR1var)

AR1var$coef[1]/sqrt(AR1var$var.coef[1,1])

plot(AR1var)

tsdisplay(AR1var$residuals)




```

Now, let's run *auto.arima*. 

``` {r auto arima}

aadelta <- auto.arima(df$delta, stepwise = F)
summary(aadelta)
plot(aadelta)
tsdisplay(aadelta$residuals)

aasign <- auto.arima(df$sign, stepwise = F)
summary(aasign)
plot(aasign)
tsdisplay(aasign$residuals)

aavar <- auto.arima((df$delta)^2, stepwise = F)
summary(aavar)
plot(aavar)
tsdisplay(aasign$residuals)

aardelta <- auto.arima(df$delta, max.q = 0, stepwise = F)
summary(aardelta)
plot(aardelta)
tsdisplay(aardelta$residuals)

aarsign <- auto.arima(df$sign, max.q = 0, stepwise = F)
summary(aarsign)
plot(aarsign)
tsdisplay(aarsign$residuals)

aarvar <- auto.arima((df$delta)^2, max.q = 0, stepwise = F)
summary(aarvar)
plot(aarvar)
tsdisplay(aarvar$residuals)


```

$$ \Delta_{t+1} = c + \beta \Delta_t + \varepsilon$$ 



``` {r AR(1) tutorial}

AR1_2 <- Arima(df$delta, order = c(1,0,0))

summary(AR1_2)

confint(AR1_2, level = 0.95)


```